The symmetry problem of the \textit{Fermi} and eROSITA bubbles: A proof-of-concept study

### 2022-02-21 paper.tex
update

\begin{abstract}
A decade ago, the \textit{Fermi Gamma-Ray Space Telescope} reveals two large bubbles in the Galaxy,\
which extend nearly symmetrically $\sim50^{\circ}$ on either side of the Galactic center (GC).\
The recent discovery of giant eROSITA bubbles also shows a symmetrically enhanced X-ray\
emission from the shells around the eROSITA bubbles.\
To this end, previous simulation-based works based on active galactic nucleus (AGN) jets model\
assumed that the jets are vertical to the Galactic disk;\
however, there is a lack of preferred orientation of jets with respect to the Galactic disk.\
Using three-dimensional (3D) special relativistic hydrodynamic (SRHD) simulations that\
include cosmic rays (CRs) and thermal gas, we show that a thin interstellar medium (ISM) disk is an essential\
component for the symmetry \textit{Fermi} and eROSITA bubbles (collectively, Galactic bubbles).\
We also produce simulated gamma-ray and\
microwave spectra from the inverse Compton scattering and synchrotron,\
and compare them with the observed counterparts in order to verify\
an oblique AGN jets model. We find that\
(1) The simulated Galactic bubbles are nearly symmetric about the GC albeit\
the bipolar jets are at an angle $45^{\circ}$ with respect to\
the rotation axis of the Galatic plane (disc normal hereafter).\
(2) The edge of the eROSITA bubbles is a forward shock,\
originally driven by oblique bipolar jets emanating from the GC 12.39 million years ago,\
and significantly stretched by the stratified atmosphere afterwards.\
(3) Followed by the forward shock is a tangling contact discontinuity,\
which we interpret as the edge of the \textit{Fermi} bubbles\
composed of turbulent and high-temperature ($\sim2$ keV)\
plasma in pressure balance with the external medium.
(4) According to the microwave and leptonic gamma-ray spectra, the best-fitting\
CRe power-law index is found to be 2.4.\
The broad agreement between simulated and observed multiwavelength features\
has thus released a caveat:\
the jets shall be vertical to the disc normal.
\end{abstract}

@@

 %originally driven by an oblique bipolar jets emanating from the GC 12.39 million years ago,\
 %and significantly stretched by the stratified atmosphere afterwards.

 % shock wave, followed by a contact discontinuity, moves to the right.

 % the innermost bubbles is an expanding reverse shock
 % the forward shock

@@

  Second, as shown in \Cref{fig__jetI5+ismSeed3-45deg},\
  the half-width of the outermost bubbles is around 7 kpc,\
  corresponding to an half angular width $\sin^{-1}(7 \text{ kpc}/R_{\odot})\sim122^{\circ}$,\
  which is as wide as the eROSITA bubbles in the simulated X-ray map (top panel in \Cref{fig__xray_0.8keV_angle_000}).\
  We therefore suggest that the nature of eROSITA bubbles is indeed a forward shocks\
  that have been driven into the northern and the southern Galactic halo,
  as previously proposed by \citet{Predehl2020}.

@@

 %We close on a speculative note that predicts an existence of a third bubbles for
 %future observations.

>>>>

\begin{abstract}
The \textit{Fermi Gamma-Ray Space Telescope} reveals two large bubbles in the Galaxy,\
which extend nearly symmetrically $\sim50^{\circ}$ on either side of the Galactic center (GC).\
The recent discovery of giant eROSITA bubbles also shows a symmetrically enhanced X-ray\
emission from the shells around the eROSITA bubbles.\
To this end, previous simulation-based works based on active galactic nucleus (AGN) jets model\
assumed that the jets are vertical to the Galactic disk;\
however, there is a lack of preferred orientation of jets with respect to the Galactic disk.\
Using three-dimensional (3D) special relativistic hydrodynamic (SRHD) simulations that\
include cosmic rays (CRs) and thermal gas, we show that a thin interstellar medium (ISM) disk is an essential\
component for the symmetry \textit{Fermi} and eROSITA bubbles (collectively, Galactic bubbles).\
We also produce simulated gamma-ray and\
microwave spectra from the inverse Compton scattering and synchrotron,\
and compare them with the observed counterparts in order to verify\
an oblique AGN jets model. We find that\
(1) The simulated Galactic bubbles are nearly symmetric about the GC albeit\
the bipolar jets are at an angle $45^{\circ}$ with respect to\
the rotation axis of the Galatic plane (disc normal hereafter).\
(2) The edge of the eROSITA bubbles is a forward shock,\
originally driven by oblique bipolar jets emanating from the GC 12.39 million years ago,\
and significantly stretched by the stratified atmosphere afterwards.\
(3) Followed by the forward shock is a tangling contact discontinuity,\
which we interpret as the edge of the \textit{Fermi} bubbles\
composed of turbulent and high-temperature ($\sim2$ keV)\
plasma in pressure balance with the external medium.
(4) According to the microwave and leptonic gamma-ray spectra, the best-fitting\
CRe power-law index is found to be 2.4.\
The broad agreement between simulated and observed multiwavelength features\
has thus released a caveat:\
the jets shall be vertical to the disc normal.
\end{abstract}

@@

@@

  Second, as shown in \Cref{fig__jetI5+ismSeed3-45deg},\
  the half-width of the outermost bubbles is around 7 kpc,\
  corresponding to an half angular width $\sin^{-1}(7 \text{ kpc}/R_{\odot})\sim122^{\circ}$,\
  which is as wide as the eROSITA bubbles in the simulated X-ray map (top panel in \Cref{fig__xray_0.8keV_angle_000}).\
  We therefore suggest that the eROSITA bubble shells are a signature of compressed forward shocks\
  that have been driven into the northern and the southern Galactic halo,
  as previously proposed by \citet{Predehl2020}.

@@
